Assessment of forestry road network projection methods using the “roads” package in R

### 2022-08-23 analysis/paper/paper.Rmd
add ref for Fmeasure etc #5

The projections were assessed for their spatial accuracy in two different ways. First, difference maps were created by subtracting the road presence rasters of the projected networks with the observed network in order to highlight discrepancies between the networks. The observed and projected road network shapefiles were then overlain on the difference maps and regions of major discrepancies were investigated using Google satellite imagery in QGIS to investigate potential causes of the discrepancy. Secondly, a matching analysis using the three binary metrics (road presence, road disturbance footprint, and forestry disturbance footprint) was performed. This matching analysis compared the three metric rasters for all projection methods with the observed metric rasters and calculated percentages of the TSA that matched and didn't match. All pixels in the Revelstoke TSA for the projected networks fell into 5 categories when compared with the observed raster: 1) pre-existing roads; 2) agree roadless, both projected and observed show no roads; 3) agree roaded, both projected and observed show new roads; 4) false positive, the projection showed a road but no roads were observed; and 5) false negative, the projection showed no roads but a road was observed. Using these 5 categories, the landscape within the Revelstoke TSA was categorized for each projection method and summarized. We also calculated the sensitivity, precision and F-measure where $sensitivity = \frac{\text{agree roaded}}{(\text{agree roaded} + \text{false negative)}}$, $precision = \frac{\text{agree roaded}}{(\text{agree roaded} + \text{false positive)}}$ and $F\text{-}measure = \frac{(2 * precision * sensitivity)}{(precision + sensitivity)}$.

@@

A mathematical weighted graph consisting of nodes and edges was used to parameterize either Kruskal's or Dijkstra's algorithms. When solving least cost paths using Dijkstra's algorithm, each node in the graph corresponded to a pixel in a raster of our study area (100m x 100m resolution). Whereas, in the graph used for solving the minimum spanning tree via Kruskal's algorithm, each node represented either a landing or a point on the existing road network. Therefore, a node can be represented by any one of three attributes in the forest road network: (i) a landing, (ii) a point on an existing road, or (iii) an intermediate point. The landing nodes represent log decks or sources of timber and the point on the existing road is the sink where the timber will end up. The intermediate points are plausible locations to construct a road and are thus free of barriers like major water bodies. [The edges connecting the nodes follow a "queens" case or allow a maximum of eight possible edges for a given node.]{.ul} Thus, our methods utilize a grid pattern which does not support curvature or switchbacks in road design as required in operational road design, thus we ascertain that our road projection methods are for strategic modelling purposes only. Edges are assigned weights which penalize the algorithm to account for various impedance and road construction costs. When solving least cost paths, a cost surface raster was used. The pixels of the cost surface raster corresponded to nodes in the weighted graph and thus the edge weight was the average cost between the two nodes or pixels. When solving the minimum spanning tree, the edge weight corresponded to the total cost of the least cost path between the two nodes.

>>>>

The projections were assessed for their spatial accuracy in two different ways. First, difference maps were created by subtracting the road presence rasters of the projected networks with the observed network in order to highlight discrepancies between the networks. The observed and projected road network shapefiles were then overlain on the difference maps and regions of major discrepancies were investigated using Google satellite imagery in QGIS to investigate potential causes of the discrepancy. Secondly, a matching analysis using the three binary metrics (road presence, road disturbance footprint, and forestry disturbance footprint) was performed. This matching analysis compared the three metric rasters for all projection methods with the observed metric rasters and calculated percentages of the TSA that matched and didn't match. All pixels in the Revelstoke TSA for the projected networks fell into 5 categories when compared with the observed raster: 1) pre-existing roads; 2) agree roadless, both projected and observed show no roads; 3) agree roaded, both projected and observed show new roads; 4) false positive, the projection showed a road but no roads were observed; and 5) false negative, the projection showed no roads but a road was observed. Using these 5 categories, the landscape within the Revelstoke TSA was categorized for each projection method and summarized. We used the performance metrics sensitivity, precision and F-measure to compare the results for each projection method [@lee2021].

@@

A mathematical weighted graph consisting of nodes and edges was used to parameterize either Kruskal's or Dijkstra's algorithms. When solving least cost paths using Dijkstra's algorithm, each node in the graph corresponded to a pixel in a raster of our study area (100m x 100m resolution). Whereas, in the graph used for solving the minimum spanning tree via Kruskal's algorithm, each node represented either a landing or a point on the existing road network. Therefore, a node can be represented by any one of three attributes in the forest road network: (i) a landing, (ii) a point on an existing road, or (iii) an intermediate point. The landing nodes represent log decks or sources of timber and the point on the existing road is the sink where the timber will end up. The intermediate points are plausible locations to construct a road and are thus free of barriers like major water bodies. [The edges connecting the nodes follow a "queens" case or allow a maximum of eight possible edges for a given node.]{.underline} Thus, our methods utilize a grid pattern which does not support curvature or switchbacks in road design as required in operational road design, thus we ascertain that our road projection methods are for strategic modelling purposes only. Edges are assigned weights which penalize the algorithm to account for various impedance and road construction costs. When solving least cost paths, a cost surface raster was used. The pixels of the cost surface raster corresponded to nodes in the weighted graph and thus the edge weight was the average cost between the two nodes or pixels. When solving the minimum spanning tree, the edge weight corresponded to the total cost of the least cost path between the two nodes.
